---
title: "R Notebook"
output: html_notebook
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or 

# Packages
_uncomment the first line if the chunk doesn't run_
```{r}
#install.packages("ggplot2")
library(readxl)
library(dplyr)
library(ggplot2)
```

# Set your working directory
I will use my output as an example of how this is useful:

1. getwd() returns your current working directory, assumed to be where you want to work on the project. i.e *"C:/Users/Sam/Desktop/R_project"*
2. setwd("C:/Users/Sam/Desktop/R_project") sets the working directory
3. Now imagine I am generating graphs and want to save them to a subfolder within my project folder. I create a subfolder named graphs and then type setwd

# Load in the dataset
```{r}
load("C:/Users/Sam/Desktop/R_project/Data/CrimeData.RData")
```

# Variable of interest selection.
We start by keeping a wide selection of potential explanatory variables. Choice of variables was based on model specifications in the existing literature around crime. 
```{r}
variables_chosen <- c("communityname:","state:","population:","PopDens:","householdsize:","PctFam2Par:","racepctblack:","racePctHisp:","agePct16t24:","perCapInc:","pctWSocSec:","pctWInvInc:","PctPopUnderPov:","PctNotHSGrad:","PctEmploy:","PolicPerPop:","PolicBudgPerPop:","ViolentCrimesPerPop:","nonViolPerPop:")
concat_dropped <- concat_test[variables_chosen]
```

# Cleaning the datatset
```{r}
is.data.frame(concat_dropped)#test dataframe ok
i=c(3:19) #Select variables to turn into numeric
concat_dropped[ , i] <- apply(concat_dropped[ , i], 2,
                    function(x) as.numeric(as.character(x)))#convert relevant variables in numeric
sapply(concat_dropped, class) #check ok
concat_wo_na= na.omit(concat_dropped)#Drop NA
concat_wi_na=concat_dropped[!complete.cases(concat_dropped),]#keep NA for statistics on drops
colnames(concat_wo_na)<-gsub(":","",colnames(concat_wo_na))# remove semicolons from names
explanatory_variables <- c("population","PopDens","householdsize","PctFam2Par","racepctblack","racePctHisp","agePct16t24","perCapInc","pctWSocSec","pctWInvInc","PctPopUnderPov","PctNotHSGrad","PctEmploy","PolicPerPop","PolicBudgPerPop")
```

# Figuring out which variables to keep. We have 15 currently, need at most 10 in my opinion. 

